---
title: "p8105_hw1_yw4662"
output: github_document
---
### Setup
```{r}
library(moderndive) # load the moderndive dataset
library(ggplot2) # load plotting
data("early_january_weather") #load the early_january_weather dataset from moderndive data
```

### Short description of the dataset
This dataset contains hourly weather observations for NYC airports over early January.

* **Time span**: from `r format(min(early_january_weather$time_hour),"%Y-%m-%d %H:%M")` to `r format(max(early_january_weather$time_hour),"%Y-%m-%d %H:%M")`.

* **Size**: `r nrow(early_january_weather)` rows $\times$ `r ncol(early_january_weather)` columns.

* **Variables**: `r paste(names(early_january_weather), collapse = ", ")`.

* **Mean air temperature**: `r round(mean(early_january_weather$temp, na.rm=TRUE),1)`°F

### Scatterplot: `temp` vs `time_hour`, colored by `humid`
```{r}
ggplot(early_january_weather,
            aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.7, size = 1.3) +
  labs(
    title = "Temperature over time, colored by relative humidity",
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity"
  )
```

### Brief description of patterns
Temperatures fluctuate in daily cycles, generally cooler overnight and warmer during daytime. Points with lower humidity (darker colors) tend to appear during cooler periods, suggesting a direct relationship between temperature and relative humidity.

```{r}
ggsave(
  filename = "temp_vs_time_hour_by_humid.png"
)
```

